---
title: "a3_task3_danielle_sclafani"
author: "Danielle Sclafani"
date: "10/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

## Part 1
### 1. Reading in sea urchin data

```{r}
sea_urchins <- read_csv(here("data", "urchins.csv")) %>%
  clean_names()
```
### 2. Comparing the size of red and purple sea urchins using graphs

### First Graph: histogram

```{r}
ggplot(data=sea_urchins, aes(x=size)) + geom_histogram(bins = 25) +
  facet_wrap(~ common_name) + #make two graphs based on common_name
  labs(x= "Sea Urchin Diameter (cm)", y= "Count of Sea Urchins",
       title = "Distribution of Sea Urchin Size by Common Name") +
  theme(plot.title = element_text(hjust = 0.5)) # centers the plot title
```

### Second Graph: Boxplot

```{r}
ggplot(data=sea_urchins, aes(x=common_name, y=size)) + geom_boxplot() +
  labs(x= "Sea Urchin Type", y= "Sea Urchin Diameter (cm)",
       title = "Distribution of Sea Urchin Size by Common Name") +
  theme(plot.title = element_text(hjust = 0.5)) # centers the plot title
```

### Creating summary statistics table for purple and red sea urchins

 > Step 1: finding the summary stastics

```{r}
sea_urchin_summary <- sea_urchins %>% 
  group_by(common_name) %>% # grouped by common name
  summarize(mean_size = mean(size), #found the mean of each type of sea urchin
            sd_size = sd(size), #found the standard deviation
            median_size = median(size),#found the median
            n= n()) #gave us the sample size 

sea_urchin_summary
```

> Step 2: Putting together the table with Kable

```{r}
summary_table <- sea_urchin_summary %>% 
  rename("Name" = common_name,
         "Mean (cm)" = mean_size,
         "Standard Deviation (cm)" = sd_size,
         "Median (cm)" = median_size,
         "Sample Size" = n
         )

kable(summary_table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F) %>% 
  add_header_above(c("Summary Statistics of Red and Purple Sea Urchins" = 5)) # add title to table, across the 5 rows
  
4.469797 - 6.776489

(6.776489/4.469797)-1

((6.776489-4.469797)/ 4.469797) *100

```



### Summary of graphs and table

> The histograms of the red and purple sea urchins show roughly normal distributed data. The spread differences between the red and purple sea urchins is obvious in both the histogram and boxplot graphs, on average red sea urchins have a larger diameter than purple sea urchins. Red urchins have a mean diameter of 6.78 cm with a standard deviation of 1.89 cm, while purple sea urchins have a mean of 4.47 cm and a standard deviation of 1.18 cm. The boxplots gives a clear visualization of the differences in mean with the middle line of the box showing the mean of each sample. The mean size of red sea urchin's diameter is 2.31 cm larger, or 51.6% larger, than the mean size of the purple sea urchin's diameter.

### Part 2: Purple urchins by site and year

```{r}

urchins_clean <- sea_urchins %>% #cleaning the data before converting the data to a date class, somehow the numbers were in a "function" class, so I had to clean the dates using mutate before I could use the mutate mdy function to make the numbers into a date class
  mutate(date_new = mdy(date))
class(urchins_clean$date)

purple_urchins <- urchins_clean %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date_new = mdy(date)) %>% 
  mutate(year = year(date_new)) %>%
  mutate(site_full = case_when( # creating a site column that writes the full names of the sites
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyoa Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"
  ))

  
```

## Creating a summary table - total number of purple urchins observed by site and year, starting from purple_urchins you created above and using the count() function to find counts of observations by group. You do not need to finalize this table (you’ll just use the output in a plot in the next step)

```{r}
purple_summary <- purple_urchins %>% 
  group_by(site_full, year) %>% 
  count(site_full) %>% 
  rename(observations = n, #renaming for the purpose of the title column names
         "Site_Name" = site_full)

kable(purple_summary) %>% 
  kable_styling(full_width = F) %>% 
  add_header_above(c("# of Purple Urchins recorded at each site by year" = 3))
```


## Graph of purple urchins at each site over time

```{r}

ggplot(data = purple_summary, aes(x = year, y=observations)) + geom_col() +
  facet_wrap(~Site_Name) +
  labs(x= "Year", y = "Number of Observations Recorded", 
       title = "Number of Purple Urchins Recorded at Each Site from 2008-2020") +
  theme(plot.title = element_text(hjust = 0.5))


```


